---
title: "Analysis of Products Sales"
subtitle: "Practicum II"
author: "Steffi Thattakunnil"
date: "Spring 2024"
output: pdf_document
---

```{r setupConnection, echo=FALSE, message=FALSE, warning=FALSE}
# Load the required libraries
library(DBI)
library(RMySQL)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(scales)

# Database connection details
db_server <- "sql5.freemysqlhosting.net"
db_name <- "sql5699914"
db_username <- "sql5699914"
db_password <- "NRLtz2sMsj"
db_port <- 3306

# Connect to the MySQL database
dbcon <- dbConnect(RMySQL::MySQL(), 
                   user=db_username, 
                   password=db_password, 
                   dbname=db_name, 
                   host=db_server, 
                   port=db_port)
```

## Top Five Sales Reps by Year

The table 'Top Five Sales Reps by Year' showcases the highest-achieving sales representatives for each year based on total sales. 
```{r topReps, echo=FALSE, message=FALSE, warning=FALSE}

# Analytical Query I:
top_reps_data <- dbGetQuery(dbcon, "
  SELECT
    year,
    repName,
    SUM(totalSoldAmount) AS totalSales
  FROM rep_facts
  GROUP BY year, repID
  ORDER BY year, totalSales DESC
")

# Use dplyr to limit to top five per year
top_reps <- top_reps_data %>%
  group_by(year) %>%
  slice_max(order_by = totalSales, n = 5) %>%
  ungroup()  %>%
  rename(Year = year, `Sales Representative` = repName, `Total Sales` = totalSales)

# Using kable to create a table
kable(top_reps, "latex", booktabs = TRUE, caption = "Top Five Sales Reps by Year") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"))

```

```{r topRepsTable, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}

# Analytical Query I - Print Table
print(top_reps)

```

\newpage
## Quarterly Sales per Product

The 'Quarterly Sales per Product' graph presents the comparison of product performance across all quarters for the given years.

```{r quarterlySales, echo=FALSE, message=FALSE, warning=FALSE}

# Analytical Query II:
quarterly_sales_data <- dbGetQuery(dbcon, "
  SELECT
    productName,
    year,
    quarter,
    SUM(totalSales) AS totalQuarterlySales
  FROM product_facts
  GROUP BY productName, year, quarter
  ORDER BY productName, year, quarter
")

```

```{r quarterlySalesTable, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}

# Analytical Query II - Print Table
print(quarterly_sales_data)

```

```{r quarterylySalesGraph, fig.width=10, fig.height=8, echo=FALSE, message=FALSE, warning=FALSE}

# Analytical Query II - Plot Graph

ggplot(quarterly_sales_data, aes(x = quarter, y = totalQuarterlySales, group = year, color = year)) +
  geom_line() +
  facet_wrap(~productName, scales = "free_y", ncol = 3) + 
  scale_y_continuous(labels = label_comma()) + # Use label_comma to avoid scientific notation
  labs(title = "Quarterly Sales per Product : ", x = "Quarter", y = "Total Sales Amount") +
  theme_minimal() +
  theme(legend.position = "top") 

```


\newpage
## Units Sold per Product per Region

The 'Units Sold per Product per Region' graph compares the unit sales across different regions, broken down by product. 

```{r unitsSold, echo=FALSE, message=FALSE, warning=FALSE}

# Analytical Query III:
units_sold_data <- dbGetQuery(dbcon, "
  SELECT
    productName,
    region,
    SUM(totalUnits) AS unitsSold
  FROM product_facts
  GROUP BY productName, region
")

```

```{r unitsSoldTable, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}

# Analytical Query III - Print Table
print(units_sold_data)

```

```{r unitsSoldGraph, echo=FALSE, message=FALSE, warning=FALSE}

# Analytical Query III - Plot Graph

ggplot(units_sold_data, aes(x = region, y = unitsSold, group = productName, color = productName)) +
  geom_line() +
  scale_y_continuous(labels = label_comma()) + # Use label_comma to avoid scientific notation
  labs(title = "Units Sold per Product per Region", x = "Region", y = "Units Sold")
```


\newpage
## Average Yearly Sales per Sales Rep

The 'Average Yearly Sales per Sales Rep' graph depicts the average sales achieved per transaction by each sales rep over the years.

```{r avgSales, echo=FALSE, message=FALSE, warning=FALSE}

# Analytical Query IV:
avg_sales_data <- dbGetQuery(dbcon, "
  SELECT
    repName,
    year,
    ROUND(SUM(totalSoldAmount) / SUM(totalTransactions)) as avgYearlySales
  FROM rep_facts 
  GROUP BY repName, year
")

```

```{r avgSalesTable, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}

# Analytical Query IV - Print Table
print(avg_sales_data)

```

```{r avgSalesGraph, echo=FALSE, message=FALSE, warning=FALSE}

# Analytical Query IV - Plot Graph

ggplot(avg_sales_data, aes(x = year, y = avgYearlySales, group = repName, color = repName)) +
  geom_line() +
  geom_point() +  # Add points to the line graph for clearer data points
  labs(title = "Average Yearly Sales per Sales Rep",
       x = "Year",
       y = "Average Sales",
       color = "Sales Rep") +
  theme_minimal() +
  theme(legend.position = "bottom")  

```



```{r disconnectDB, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}

# Close the database connection
dbDisconnect(dbcon)

```